Annotated example of index standardization with VAST

### 2021-12-09 doc/appendix-pcod-VAST-tweedie.Rmd
init Tweedie version

For this example, we will include a factor predictor that represents the mean estimate for each time slice. Settings used for index standardization are specified partially by specifying `purpose = "index2"` but we also explicitly provide arguments for these and other key settings here.

```{r config, echo = TRUE}
FieldConfig <- matrix(c("IID", "IID", "IID", "IID", "IID", "IID"),
  ncol = 2, nrow = 3,
  dimnames = list(
    c("Omega", "Epsilon", "Beta"),
    c("Component_1", "Component_2")
  )
)
FieldConfig
RhoConfig <- c("Beta1" = 0, "Beta2" = 0, "Epsilon1" = 0, "Epsilon2" = 0)
```

@@

settings <- make_settings(
  n_x = 205, # number of vertices in the SPDE mesh
  Region = "User",
  purpose = "index2", # index of abundance with Gamma for positive catches
  fine_scale = TRUE, # use bilinear interpolation from the INLA 'A' matrix
  zone = 9,
  FieldConfig = FieldConfig,
  RhoConfig = RhoConfig,
  ObsModel = c(2, 0), # this is for conventional logit-linked delta-Gamma; c(10, 2) for Tweedie
  bias.correct = FALSE,
  use_anisotropy = FALSE,
  max_cells = Inf, # use all grid cells from the extrapolation grid
  knot_method = "grid" # or "samples"
)
```

@@

```{r fit, echo = TRUE, results = 'hide'}
 # create folder for saved output:
dir.create(paste0(here("doc", "appendix-VAST")), showWarnings = FALSE)

# effort is 1 when using CPUE instead of observed weight as the response:
pcod$effort <- 1
pcod <- as.data.frame(pcod) # ensure not a tibble
fit <- fit_model(
  settings = settings,
  Lat_i = pcod[, "lat"],
  Lon_i = pcod[, "lon"],
  t_i = pcod[, "year"],
  b_i = pcod[, "density"],
  a_i = pcod[, "effort"],
  input_grid = input_grid,
  working_dir = paste0(here("doc", "appendix-VAST"), "/")
)
```

@@

```{r save-plots, echo = TRUE, warning = FALSE}
plot(
  fit,
  check_residuals = FALSE,
  working_dir = paste0(here("doc", "appendix-VAST"), "/")
)
```

@@

```{r plot-knots, echo = TRUE}
knitr::include_graphics(here("doc", "appendix-VAST", "Data_and_knots.png"))
```

@@

```{r plot-density, echo = TRUE}
knitr::include_graphics(here("doc", "appendix-VAST", "ln_density-predicted.png"))
```

@@

```{r plot-index, echo = TRUE, out.width="3in"}
if (file.exists(here("doc", "appendix-VAST", "Index-Biomass.png"))) {
  knitr::include_graphics(here("doc", "appendix-VAST", "Index-Biomass.png"))
}
if (file.exists(here("doc", "appendix-VAST", "Index.png"))) {
  knitr::include_graphics(here("doc", "appendix-VAST", "Index.png"))
}
```

>>>>

For this example, we will include a factor predictor that represents the mean estimate for each time slice. Settings used for index standardization are specified partially by defining `purpose = "index2"` but we also explicitly provide arguments for these and other key settings here. Specifically, we set the elements of `RhoConfig` to 0 to specify that we want the intercepts and spatiotemporal fields to be independent (which is the default).

VAST is constructed to accommodate many different model structures, including delta or hurdle models, and thus requires the user to specify arguments for the spatial and spatiotemporal fields of two model components. In this case, we want to fit a model with a single component, using the Tweedie distribution. To do so, we must turn off estimation of omega and epsilon for 1st linear predictor by entering 0 for those elements in the object we'll use to define the configuration of these fields `FieldConfig`.

```{r config, echo = TRUE}
FieldConfig <- matrix(c("0", "0", "IID", "Identity", "IID", "IID", "IID", "Identity"),
  ncol = 2, nrow = 4,
  dimnames = list(
    c("Omega", "Epsilon", "Beta", "Epsilon_year"),
    c("Component_1", "Component_2")
  )
)
FieldConfig

RhoConfig <- c("Beta1" = 0, "Beta2" = 0, "Epsilon1" = 0, "Epsilon2" = 0)
```

@@

settings <- make_settings(
  n_x = 205, # number of vertices in the SPDE mesh
  Region = "User",
  purpose = "index2", # use recommended defaults for an index of abundance
  fine_scale = TRUE, # use bilinear interpolation from the INLA 'A' matrix
  zone = 9,
  FieldConfig = FieldConfig,
  RhoConfig = RhoConfig,
  ObsModel = c(10, 2), # use the Tweedie distribution as the observation model
  bias.correct = FALSE,
  use_anisotropy = FALSE,
  max_cells = Inf, # use all grid cells from the extrapolation grid
  knot_method = "grid" # or "samples"
)
```

@@

```{r fit, echo = TRUE, results = 'hide'}
 # create folder for saved output:
dir.create(paste0(here("doc", "appendix-VAST-tweedie")), showWarnings = FALSE)

# effort is 1 when using CPUE instead of observed weight as the response:
pcod$effort <- 1
pcod <- as.data.frame(pcod) # ensure not a tibble
fit <- fit_model(
  settings = settings,
  Lat_i = pcod[, "lat"],
  Lon_i = pcod[, "lon"],
  t_i = pcod[, "year"],
  b_i = pcod[, "density"],
  a_i = pcod[, "effort"],
  input_grid = input_grid,
  working_dir = paste0(here("doc", "appendix-VAST-tweedie"), "/")
)
```

@@

```{r save-plots, echo = TRUE, warning = FALSE}
plot(
  fit,
  check_residuals = FALSE,
  working_dir = paste0(here("doc", "appendix-VAST-tweedie"), "/")
)
```

@@

```{r plot-knots, echo = TRUE}
knitr::include_graphics(here("doc", "appendix-VAST-tweedie", "Data_and_knots.png"))
```

@@

```{r plot-density, echo = TRUE}
knitr::include_graphics(here("doc", "appendix-VAST-tweedie", "ln_density-predicted.png"))
```

@@

```{r plot-index, echo = TRUE, out.width="3in"}
if (file.exists(here("doc", "appendix-VAST-tweedie", "Index-Biomass.png"))) {
  knitr::include_graphics(here("doc", "appendix-VAST-tweedie", "Index-Biomass.png"))
}
if (file.exists(here("doc", "appendix-VAST-tweedie", "Index.png"))) {
  knitr::include_graphics(here("doc", "appendix-VAST-tweedie", "Index.png"))
}
```
